### 2015-07-01 Harris_2015/manuscript-materials/draft.Rmd
more dissertation stuff

Inferring species interactions from observational data is one of the most
controversial tasks in community ecology. One difficulty is that a single
pairwise interaction can ripple through an ecological network and produce
surprising indirect consequences. For example, two competing species would
ordinarily correlate negatively in space, but this effect can be reversed in
the presence of a third species that is capable of outcompeting both of them
when it is present. Here, I apply models from statistical physics, called
Markov networks or Markov random fields, that can predict the direct and
indirect consequences of any possible species interaction matrix. Interactions
in these models can be estimated from observational data via maximum
likelihood. Using simulated landscapes with known pairwise interaction
strengths, I evaluated Markov networks and several existing approaches. The
Markov networks consistently outperformed other methods, correctly isolating
direct interactions between species pairs even when indirect interactions or
abiotic environmental effects largely overpowered them. A linear
approximation, based on partial covariances, also performed well as long as
the number of sampled locations exceeded the number of species in the data.
Indirect effects reliably caused a common null modeling approach to produce
incorrect inferences, however.

@@

***Conditional relationships and Markov networks.***
Ecologists are often interested in inferring direct interactions between
species, controlling for the indirect influence of other species. In
statistical terms, this implies that ecologists want to estimate *conditional*
("all-else-equal") relationships, rather than *marginal* ("overall") relationships. 
The most familiar conditional relationship
is the partial correlation, which indicates the portion of the sample correlation
between two species that remains after controlling for other variables in the
data set [@albrecht_spatial_2001]. The example with the shrubs and trees in
Figure 1 shows how the two correlation measures can have opposite signs, 
and suggests that the partial correlation is more relevant
for drawing inferences about species interactions (e.g. competition). 
Markov networks extend this approach to non-Gaussian data, much as
generalized linear models do for linear regression [@lee_learning_2012].

@@

Estimating the marginal relationships predicted by a Markov network is more
difficult than estimating conditional relationships, because doing so requires
absolute probability estimates. Turning the relative probability given by
Equation 1 into an absolute probability entails scaling by a *partition
function*, $Z(\alpha, \beta)$, which is defined so that that the probabilities of all
possible assemblages that could be produced by the model sum to one (bottom of
Figure 2B). Calculating $Z(\alpha, \beta)$ exactly, as is done in this paper,
quickly becomes infeasible as the number of species increases: with
$2^N$ possible assemblages of $N$ species, the number of bookkeeping operations
required for exact inference spirals exponentially into the billions and beyond.
Numerous techniques are available for working with Markov networks that keep
the computations tractable, e.g. via analytic approximations
[@lee_learning_2012] or Monte Carlo sampling [@salakhutdinov_learning_2008],
but they are beyond the scope of this paper.

***Simulations.***
In order to compare different methods for drawing inferences from
observational data, I simulated two sets of landscapes using known
parameters.

@@

***Other inference techniques for comparison.***
After fitting Markov networks to the simulated landscapes described above, I
used several other techniques for inferring the sign and strength of marginal
associations between pairs of species (Appendix B).

>>>>

Estimating species interactions from observational data is one of the most
controversial tasks in community ecology. One difficulty is that a single
pairwise interaction can ripple through an ecological network and produce
surprising indirect consequences. For example, two competing species would
ordinarily correlate negatively in space, but this effect can be reversed in
the presence of a third species that is capable of outcompeting both of them
when it is present. Here, I apply models from statistical physics, called
Markov networks or Markov random fields, that can predict the direct and
indirect consequences of any possible species interaction matrix. Interactions
in these models can be estimated from observational data via maximum
likelihood. Using simulated landscapes with known pairwise interaction
strengths, I evaluated Markov networks and several existing approaches. The
Markov networks consistently outperformed other methods, correctly isolating
direct interactions between species pairs even when indirect interactions or
abiotic environmental effects largely overpowered them. A linear
approximation, based on partial covariances, also performed well as long as
the number of sampled locations exceeded the number of species in the data.
Indirect effects reliably caused a common null modeling approach to produce
incorrect inferences, however.

@@

***Conditional relationships and Markov networks.***
Ecologists are often interested in learning about direct interactions between
species, controlling for the indirect influence of other species. In
statistical terms, this implies that ecologists want to estimate *conditional*
("all-else-equal") relationships, rather than *marginal* ("overall") relationships.
The most familiar conditional relationship
is the partial correlation, which indicates the portion of the sample correlation
between two species that remains after controlling for other variables in the
data set [@albrecht_spatial_2001]. The example with the shrubs and trees in
Figure 1 shows how the two correlation measures can have opposite signs,
and suggests that the partial correlation is more relevant
for drawing inferences about species interactions (e.g. competition).
Markov networks extend this approach to non-Gaussian data, much as
generalized linear models do for linear regression [@lee_learning_2012].

@@

Estimating the marginal relationships predicted by a Markov network is more
difficult than estimating conditional relationships, because doing so requires
absolute probability estimates. Turning the relative probability given by
Equation 1 into an absolute probability entails scaling by a *partition
function*, $Z(\alpha, \beta)$, which is defined so that that the probabilities of all
possible assemblages that could be produced by the model sum to one (bottom of
Figure 2B). Calculating $Z(\alpha, \beta)$ exactly, as is done in this paper,
quickly becomes infeasible as the number of species increases: with
$2^N$ possible assemblages of $N$ species, the number of bookkeeping operations
required for exact estimation spirals exponentially into the billions and beyond.
Numerous techniques are available for working with Markov networks that keep
the computations tractable, e.g. via analytic approximations
[@lee_learning_2012] or Monte Carlo sampling [@salakhutdinov_learning_2008],
but they are beyond the scope of this paper.

***Simulations.***
In order to compare different methods, I simulated two sets of landscapes using known
parameters.

@@

***Comparison with existing techniques.***
After fitting Markov networks to the simulated landscapes described above, I
used several other techniques for determining the sign and strength of marginal
associations between pairs of species (Appendix B).

### 2015-07-01 stochastic/chapter-3.Rmd
stochastic stuff
more dissertation stuff

@@

To the extent that species interactions are important for community assembly, ecologists generally expect them to leave a signature on species' co-occurrence patterns.  But using that signature to infer the underlying species interactions has been much harder.  Disagreements about how to draw these inferences led to an acrimonious, decade-long argument among community ecologists [@lewin_santa_1983], where essentially all of the proposed statistical approaches were criticized for having high error rates, a poor match with the underlying ecological questions, computational infeasibility, or all three [@connor_assembly_1979; @strong_ecological_1984; @hastings_can_1987]. [[**The next two sentences are unnecessarily confrontational and should probably be removed**]] *Despite these difficulties, ecologists have considered the underlying questions so important that, for the last few decades, they have continued to address them with whatever methods were available, even when these methods are known to perform poorly. For example, the False Discovery Rates above 95% demonstrated in @gotelli_empirical_2009 have not prevented wide adoption of method proposed in that paper*

During the time that ecologists have struggled to identify the interactions among dozens of species, bioinformaticians have largely solved the analogous problem of estimating interactions among *thousands* of genes from their co-expression levels (e.g. @friedman_sparse_2008). The discrepancy is due to the fact that ecologists have focused, almost exclusively, on overall co-occurrence rates [@connor_assembly_1979; @gotelli_empirical_2009; @gilpin_factors_1982; @veech_probabilistic_2013], and most of the test statistics in common use are actually minor variations on the sample correlation [*evidence to be added to Chapter 2*]. As ecologists know, however, the correlation between two species will often reflect other factors beyond their direct pairwise interactions (Figure 1, *to be added*). While some narrowly-focused methods have been proposed for incorporating a small number of specific factors into ecologists' null models at a time [@lessard_strong_2011], we still don't have a good way to scale these approaches up for cases with many factors acting simultaneously.

In other fields, scientists have a number of procedures for controlling for the influence of extraneous factors and focusing on the direct relationship between a single pair of variables.  The most familiar of these is the partial correlation. Rather than describing the *overall* relationship between two variables across *all* conditions, partial correlations (like regression coefficients) describe the portion of the relationship that remains after the other variables in the data have been accounted for.  @harris_estimating_2015 showed that the partial correlation can do a surprisingly good job of extracting accurate information about pairwise species interactions from co-occurrence data, and that Markov networks (undirected graphical models also known as Markov random fields), which do not assume normally-distributed observations do even better (cf @azaele_inferring_2010 and @fort_statistical_2013). A Markov network defines a probability distribution over possible binary species assemblages, and its coefficients (including one coefficient describing the conditional relationship between each pair of species in the network) can be estimated by maximum likelihood.

Unfortunately, Markov networks have an intractable likelihood function whose computational complexity more than doubles each time a new species is added to the model. Thus, while Harris (2015) was able to use off-the-shelf optimization techniques to maximize the likelihood for networks of 20 species, extending the same approach to networks of 50-100 species would be completely infeasible. If ecologists want the model to account for abiotic variation among sites on the landscape, then these expensive computations need to be repeated independently for every site, increasing the computational burden even further.

In this paper, I present a different way of optimizing the likelihood, called *stochastic approximation* [@robbins_stochastic_1951; @salakhutdinov_efficient_2012], which replaces the intractable computations with tractable Monte Carlo estimates of the same quantities. Despite the sampling error introduced by this substitution, stochastic approximation provides strong guarantees for eventual convergence to the maximum likelihood estimate. This change in approach makes it feasible to estimate the interactions among hundreds of species from observational data while also accounting for possible responses to abiotic factors.

@@

As discussed in @azaele_inferring_2010, @fort_statistical_2013, and @harris_estimating_2015, Markov networks such as the Ising model [@cipra_introduction_1987] can be used to describe community structure as follows.  Each species is represented by a binary random variable, describing whether it is present or absent.  A species’ conditional probability of presence under a given set of biotic and abiotic conditions depends on its coefficients. The coefficients linking species to one another describe their conditional associations: *all else equal*, a species will occur less often in the presence of its competitors and other species with which it is negatively associated, and more often in the presence of its mutualists.  These conditional probabilities can be combined to calculate the probability of observing any given combination of presences and absences.  For example, hypothetical assemblages including many pairs of competitively exclusive species at the same site will tend to be very improbable under the model.

Unfortunately, the extremely large number of possible assemblages makes this joint probability intractable when the number of species is large.  However, it is generally straightforward to simulate examples of assemblages that are consistent with graphical models such as these, using Monte Carlo techniques like Gibbs sampling [@salakhutdinov_efficient_2012]. Gibbs sampling is especially convenient for these models because it only requires the conditional probabilities, which are straightforward to compute (Harris 2015; Appendix).

In each of the scenarios below, the “observed” landscapes were simulated using Gibbs sampling from a pre-specified set of “true” parameters.

@@

In order to evaluate partial correlations' ability to recover the "true" parameters that generated a presence-absence matrix of interest, I first needed to generate such matrices from known processes.

To demonstrate that the stochastic approximation method worked, I first simulated 10 landscapes with 20 species and 500 sites, using the Gibbs sampling approach described in @harris_estimating_2015.  These landscapes had few enough species that the `rosalia` package for estimating Markov networks [@harris_rosalia_2015] could find the penalized maximum likelihood estimates for the Markov network in a reasonable amount of time.  For each of these simulated landscapes, I then fit the same model using both the exact approach from @harris_estimating_2015 and the stochastic approximation method described below.

@@

A Markov network describes a probability distribution over possible assemblages, and this distribution can be summarized without loss of information by its *sufficient statistics*.  For the model presented here, the sufficient statistics include: the number of occurrences for each species, the number of co-occurrences between each species, and the cross-product between the species and environment matrices [[*Appendix, not yet written*]].  One way to look at maximum likelihood estimation is that it minimizes the discrepancy between the sufficient statistics observed in the data and the average sufficient statistics produced by samples from the model [[ref?]]. Because fully-observed Markov networks, such as the ones analyzed here, have unimodal likelihood functions [@murphy_machine_2012], any iterative procedure that can reduce this discrepancy quickly enough will converge to the global maximum likelihood estimate.

The `rosalia` package reduces the discrepancy between observed and predicted sufficient statistics by calculating it exactly, averaging over *all* possible presence-absence combinations, but this approach is intractable for larger networks. Stochastic approximation [refs] is a procedure that allows us to approximate the expected values of the sufficient statistics by averaging over a more manageable number of simulated assemblages during each model-fitting iteration, while still retaining maximum likelihood convergence guarantees.  The procedure iterates through the following three steps as many times as needed (5000 for these analyses):

The simulations in Step 1 used Gibbs sampling to generate examples of landscapes based on the model’s current parameter estimates. While the simulated landscapes produced by Gibbs sampling are serially autocorrelated, statisticians have shown that this difficulty does not prevent convergence to the maximum likelihood estimate [@younes_convergence_1999; @salakhutdinov_efficient_2012].

The step size parameter in Step 3, $\alpha_t$, decreased after each iteration according to a generalized harmonic sequence, $\alpha_t = \alpha_0 500/(500 - t - 1)$, which satisfies the criteria in @powell_approximate_2007. $\alpha_0$ was 1.0 for species’ intercepts and environmental responses, and one tenth as large for their pairwise interaction coefficients.

@@

The squared deviations between the exact estimates produced by the `rosalia` package and the ones produced by stochastic approximation quickly decayed to negligible levels [[Figure XA]], indicating that the stochastic approximation procedure was implemented correctly and worked as the mathematical theory predicts.

@@

For the larger landscape, I found that the stochastic approximation approach achieved reasonably good performance after less than a minute of optimization [Figure XB]. With additional time, was able to recover most of the variance in species’ pairwise interactions from observational data ($R^2 = 0.66$; Figure YA), and nearly all of the variance in their responses to environmental variables ($R^2 = 0.95$, not shown).

While the maximum likelihood estimates produced by the Markov network worked well, the sample correlation, explained a negligible fraction of the variance in species’ pairwise interactions ($R^2 = 0.03$, Figure YB).  To the extent that ecologists’ other test statistics for co-occurrence are similar to the sample correlation [[*evidence to be added to chapter 2*]], they can be expected to perform equally poorly. As shown previously, the partial correlation was a much better indicator of species’ true interactions than the sample correlation ($R^2 = 0.44$ Figure YC).

@@

For decades, ecologists have relied on minor variations on the correlation coefficient as their test statistics for inferring species interactions from observational data @harris_estimating_2015. As shown here and in @harris_estimating_2015, however, these inferences can only be made reliably by methods that control for other species in the network, such as partial correlations or Markov networks.  The biggest computational problem with large Markov networks is the impossibility of averaging over all their possible states, but these results demonstrate that this step can be avoided.  Stochastic approximation is able to recover the same estimates as exact methods while scaling gracefully to networks with hundreds of species.

>>>>

# Abstract

Species interactions are believed to play an important role in community structure, but detecting their influence in observed communities has stymied ecologists for decades.  In recent years, several papers have suggested that ecologists could solve this long-standing problem by fitting Markov networks (undirected graphical models also known as Markov random fields) to the data by maximum likelihood or similar methods.  This approach allows ecologists to control for extraneous factors when estimating the direct interactions between a given pair of species, and previous work has shown that this kind of control is required for reliable inferences.  Unfortunately, the likelihood function for Markov networks quickly becomes intractable once more than about 20 species are included in the model, due to the exponentially large number of possible combinations of presences and absences. Additionally, the methods that have been proposed so far do not account for environmental heterogeneity, and attempt to explain *all* of the co-occurrence patterns in ecological data sets with species interactions. In this paper, I introduce *stochastic approximation* as a method for fitting these models to large ecological networks where hundreds of species and multiple abiotic factors can all affect a given species’ occurrence probability.  This approach, which converges to the maximum likelihood estimate with probability one, greatly expands the range of data sets we can analyze and the number of ecological questions we can address with these methods.

@@

To the extent that species interactions are important for community assembly, ecologists generally expect them to leave a signature on species' co-occurrence patterns.  But using that signature to infer the underlying species interactions has been much harder.  Disagreements about how to draw these inferences led to an acrimonious, decade-long argument among community ecologists [@lewin_santa_1983], where essentially all of the proposed statistical approaches were criticized for having high error rates, a poor match with the underlying ecological questions, computational infeasibility, or all three [@connor_assembly_1979; @strong_ecological_1984; @hastings_can_1987]. As documented below, ecologists have, for the most part, not solved these problems in the ensuing decades.

During the time that ecologists have struggled to identify the interactions among dozens of species, bioinformaticians have largely solved the analogous problem of estimating interactions among *thousands* of genes from their co-expression levels (e.g. @friedman_sparse_2008). The discrepancy is due to the fact that ecologists have focused, almost exclusively, on overall co-occurrence rates [@connor_assembly_1979; @gotelli_empirical_2009; @gilpin_factors_1982; @veech_probabilistic_2013] or on closely-related values such as correlation coefficients [@pollock_understanding_2014; see @faisal_inferring_2010 for a recent exception]. As ecologists know, however, the correlation between two species will often reflect other factors beyond their direct pairwise interactions (Figure 1). When these factors are not accounted for, co-occurrence rates can reliably lead to incorrect inferences [@harris_estimating_2015]. While some methods have been proposed for incorporating a small number of specific factors (such as geographic or environmental dissimilarity) into ecologists' null models at a time [@lessard_strong_2011], we still don't have a good way to scale these approaches up for cases with many biotic and abiotic factors acting simultaneously.

A number of procedures have been developed for controlling for the influence of extraneous factors and focusing on the direct relationship between a single pair of variables, i.e. for estimating the *conditional* relationship between them.  The most familiar of these is the partial correlation. Rather than describing the *overall* relationship between two variables across *all* conditions, partial correlations (like regression coefficients) describe the portion of the relationship that remains after the other variables in the data have been accounted for.  @harris_estimating_2015 found that the partial correlation can do a surprisingly good job of extracting accurate information about pairwise species interactions from binary co-occurrence data in some circumstances, but the fact that this approach assumes multivariate Gaussian data makes it less appealing.

Recently, several papers have suggested that ecologists could use Markov networks (undirected graphical models also known as Markov random fields) to estimate conditional relationships from binary data by maximum entropy or maximum likelihood (@azaele_inferring_2010; @fort_statistical_2013; @harris_estimating_2015). A Markov network defines a probability distribution over possible binary species assemblages, and its coefficients (including one coefficient describing the conditional relationship between each pair of species in the network).  Unfortunately, Markov networks have an intractable likelihood function whose computational complexity more than doubles each time a new species is added to the model. Thus, while @azaele_inferring_2010 and @harris_estimating_2015 were able to optimize an exact objective function to maximize the constrained likelihood for networks of up to 20 species, extending these approaches to networks of 50-100 species would be completely infeasible. If ecologists want the model to account for abiotic variation among sites on the landscape, then these expensive computations need to be repeated independently for every site, increasing the computational burden even further.

In this paper, I present a different way of optimizing the likelihood, called *stochastic approximation* [@robbins_stochastic_1951; @salakhutdinov_efficient_2012], which replaces the intractable computations with tractable Monte Carlo estimates of the same quantities. Despite the sampling error introduced by this substitution, stochastic approximation provides strong guarantees for eventual convergence to the maximum likelihood estimate. This change in approach makes it feasible to estimate the interactions among hundreds of species from observational data while also accounting for possible responses to abiotic factors, possibly giving ecologists some power to distinguish between species pairs that co-occur due to shared environmental tolerances from those that co-occur due to direct interactions like mutualism.

@@

As discussed in @azaele_inferring_2010, @fort_statistical_2013, and @harris_estimating_2015, Markov networks such as the Ising model [@cipra_introduction_1987] can be used to describe community structure as follows.  Each species is represented by a binary random variable, describing whether it is present or absent.  A species’ conditional probability of presence under a given set of biotic and abiotic conditions depends on its coefficients. The coefficients linking species to one another describe their conditional associations: *all else equal*, a species will occur less often in the presence of its competitors and other species with which it is negatively associated, and more often in the presence of its mutualists.

[[math for conditional distributions]]

These conditional probabilities can be combined to calculate the probability of observing any given combination of presences and absences [[Appendix?]].  For example, hypothetical assemblages including many pairs of competitively exclusive species at the same site will tend to be very improbable under the model.  Unfortunately, the extremely large number of possible assemblages makes this joint probability intractable when the number of species is large.  However, it is generally straightforward to simulate examples of assemblages that are consistent with graphical models such as these, using Monte Carlo techniques like Gibbs sampling [@salakhutdinov_efficient_2012]. Gibbs sampling is especially convenient for these models because it only requires the conditional probabilities, which are straightforward to compute (Harris 2015; Appendix).

In each of the scenarios below, the “observed” landscapes were simulated using Gibbs sampling from a pre-specified set of “true” parameters, then stochastic approximation was used to recover these parameters from the simulated presence-absence data.

A Markov network describes a probability distribution over possible assemblages, and this distribution can be summarized without loss of information by its *sufficient statistics*.  For the model presented here, the sufficient statistics include: the number of occurrences for each species, the number of co-occurrences between each species, and the cross-product between the species and environment matrices [[*Appendix, not yet written*]].  One way to look at maximum likelihood estimation is that it minimizes the discrepancy between the sufficient statistics observed in the data and the average sufficient statistics produced by samples from the model [[ref?]]. Because fully-observed Markov networks, such as the ones analyzed here, have unimodal likelihood functions [@murphy_machine_2012], it is possible to find the global optimum by climbing the likelihood surface. At this point, any discrepancies between the sufficient statistics of the model and of the data will have been eliminated.

In order to reduce the discrepancy between the observed and predicted sufficient statistics, the `rosalia` calculates this different exactly, averaging over *all* possible presence-absence combinations. This approach is intractable for larger networks, as the number of possibilities grows exponentially with the size of the network. Stochastic approximation [refs] is a procedure that allows us to approximate the expected values of the sufficient statistics by averaging over a more manageable number of simulated assemblages during each model-fitting iteration, while still retaining maximum likelihood convergence guarantees [refs].  The procedure iterates through the following three steps as many times as needed (5000 for these analyses):

The simulations in Step 1 used Gibbs sampling to generate examples of landscapes based on the model’s current parameter estimates. While the simulated landscapes produced by Gibbs sampling are serially autocorrelated, statisticians have shown that this merely slows convergence to the maximum likelihood estimate rather than preventing it altogether [@younes_convergence_1999; @salakhutdinov_efficient_2012].

The step size parameter in Step 3, $\alpha_t$, decreased after each iteration according to a generalized harmonic sequence, $\alpha_t = \alpha_0 500/(500 - t - 1)$, which satisfies the criteria in @powell_approximate_2007. $\alpha_0$ was 1.0 for species’ intercepts and environmental responses. Because of the enormous number of pairwise interaction coefficients, I set $\alpha_0$ to 0.1 for the $\beta$ parameters.

@@

In order to assess the models' ability to recover the "true" parameters that generated a presence-absence matrix of interest, I first needed to generate such matrices from known processes.

To demonstrate that my stochastic approximation implementation could converge to the global optimum, I first simulated 10 landscapes with 20 species and 500 sites, using the Gibbs sampling approach described in @harris_estimating_2015.  These landscapes had few enough species that the `rosalia` package for estimating Markov networks [@alia_2015] could find the penalized maximum likelihood estimates for the Markov network in a reasonable amount of time.  For each of these simulated landscapes, I then fit the same model using both the exact approach from @harris_estimating_2015 and the stochastic approximation method described above.

@@

## Comparison with directed models

[[In response to Chapter 2, reviewers suggested comparisons with two other methods based on directed acyclic graphs.  The first one has a latent multivariate Gaussian, and relies on the correlation structure of that Gaussian to help determine which species interact.  The second one is based on sparse directed graphs among the species in the network. Should I address one or both of these approaches in this chapter as well?]]

@@

For the smaller communities, the squared deviations between the exact estimates produced by the `rosalia` package and the ones produced by stochastic approximation quickly decayed to negligible levels [[Figure 2A]], indicating that the stochastic approximation procedure was implemented correctly and worked as the mathematical theory predicts.

@@

For the larger landscape, I found that the stochastic approximation approach achieved reasonably good performance after less than a minute of optimization [Figure 2B]. With additional time, was able to recover most of the variance in species’ pairwise interactions from observational data ($R^2 = 0.66$; Figure 3A), and nearly all of the variance in their responses to environmental variables ($R^2 = 0.95$, not shown).

While the maximum likelihood estimates produced by the Markov network worked well, the sample correlation, explained a negligible fraction of the variance in species’ pairwise interactions ($R^2 = 0.03$, Figure 3B).  To the extent that ecologists’ other test statistics for co-occurrence are similar to the sample correlation [[*evidence to be added to chapter 2*]], they can be expected to perform equally poorly. As shown previously, the partial correlation was a much better indicator of species’ true interactions than the sample correlation ($R^2 = 0.44$ Figure 3C).

@@

For decades, ecologists have relied on poor test statistics for inferring species interactions from observational data [@harris_estimating_2015]. As shown here and in @harris_estimating_2015, however, these inferences can only be made reliably by methods that control for other species in the network, such as partial correlations or Markov networks.  The biggest computational problem with large Markov networks is the impossibility of averaging over all their possible states, but these results demonstrate that this step can be avoided in ecological analyses.  Stochastic approximation is able to recover the same estimates as exact methods for smaller networks, while scaling gracefully to networks with hundreds of species.
